---
title: "Mathy"
author: "Gilles Mathy"
format: html
editor: visual
---

https://github.com/GillouMa/Gilles-Mathy

```{r}
library(here)
library(readr)
```

```{r}
here::i_am("Gilles-Mathy.Rproj")
```
### Question 1
```{r}
courses <- read_csv(here("courses.csv"))
```

### Question 2

```{r}
library(dplyr)
library(knitr)

courses |>
  arrange(course) |> 
  rename(
    "course name"      = course,
    "identifier"       = course_id,
    "trimester"        = trimester,
    "number of grades" = nb_grades
  ) |>
  kable()

```

## 1.2 Students

### Question 3

```{r}
students <- read_csv(
  here("students.csv"),
  col_types = cols(
    id         = col_integer(),
    group      = col_factor(),
    birth_date = col_date(format = "%Y-%m-%d"),
    sex        = col_factor()
  )
)
```

The `birth_date` column of the `students` data frame is of class `r class(students$birth_date)`.
The `sex` column of the `students` data frame is of class `r class(students$sex)`

## 1.3 Grades

### Question 4 

```{r}
grades <- read_delim(
  here("grades.csv"),
  delim = ";",
  na = "unknown",
  col_types = cols(
    id        = col_integer(),
    course_id = col_integer(),
    grade     = col_double()
  ),
  show_col_types = FALSE
)
```

# Student Population Analysis

### Question 5

```{r}
library(ggplot2)

students <- students |>
  mutate(
    group = as.integer(as.character(group))   
  )

students_per_group <- students |>
  count(group)

ggplot(students_per_group, aes(x = factor(group), y = n)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Number of Students per Group",
    x = "Group",
    y = "Number of Students"
  ) +
  theme_minimal()
```

### Question 6

```{r}
gender_group <- students |>
  count(group, sex)

ggplot(gender_group, aes(x = factor(group), y = n, fill = sex)) +
  geom_col(position = "fill") +
  labs(
    title = "Gender Balance per Group (Proportion)",
    x = "Group",
    y = "Proportion",
    fill = "Sex"
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_minimal()
```

### Question 7

```{r}
library(lubridate)

students <- students |>
  mutate(
    age = round(time_length(today() - birth_date, unit = "year"))
  )

ggplot(students, aes(x = age)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of Student Ages",
    x = "Age (years)",
    y = "Number of Students"
  ) +
  theme_minimal()
```

### Question 8

```{r}
median_age_group <- students |>
  group_by(group) |>
  summarise(
    median_age = round(median(age), 0)
  ) |>
  arrange(group)

kable(median_age_group, caption = "Median Age of Students per Group")
```

### Question 9

```{r}
oldest_per_group <- students |>
  group_by(group) |>
  slice_max(age, n = 1, with_ties = FALSE) |>   # pick the oldest student
  ungroup() |>
  select(group, id, sex, age) |>                # keep only needed columns
  arrange(desc(age))                            # sort by age decreasing

kable(oldest_per_group, caption = "Oldest Student in Each Group")
```

# 3 Simple Grade Analysis

### Question 10 

```{r}
n_grades <- sum(!is.na(grades$grade))
```

The data set contains `r n_grades` grades.

### Question 11

```{r}
crypto_id <- courses |>
  filter(course == "Cryptography and Codebreaking") |>
  pull(course_id)

crypto_group_avg <- grades |>
  filter(course_id == crypto_id, !is.na(grade)) |>   
  left_join(students, by = "id") |>                  
  group_by(group) |>
  summarise(avg_crypto = mean(grade)) |>
  arrange(group)

ggplot(crypto_group_avg, aes(x = factor(group), y = avg_crypto)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Average Grade in Cryptography and Codebreaking by Group",
    x = "Group",
    y = "Average Grade"
  ) +
  theme_minimal()
```

### Question 12

```{r}
grades_trimester <- grades |>
  left_join(courses, by = "course_id") |> 
  filter(!is.na(grade))  

ggplot(grades_trimester, aes(x = grade, color = factor(trimester), fill = factor(trimester))) +
  geom_density(alpha = 0.3) +
  labs(
    title = "Distribution of Grades by Trimester",
    x = "Grade",
    y = "Density",
    color = "Trimester",
    fill  = "Trimester"
  ) +
  theme_minimal()
```

# Attendance Analysis

### Question 13

```{r}
grades_per_student <- grades |>
  filter(!is.na(grade)) |>            
  count(id, name = "n_grades") |>     
  left_join(students, by = "id") |>   
  select(id, group, sex, n_grades) |> 
  arrange(id)

grades_per_student |>
  slice_head(n = 10) |>
  knitr::kable(caption = "First 10 Students: Number of Grades per Student")
```

```{r}
summary_stats <- grades_per_student |>
  summarise(
    minimum = min(n_grades),
    maximum = max(n_grades),
    average = mean(n_grades),
    median  = median(n_grades)
  )

knitr::kable(summary_stats, caption = "Summary Statistics of Grades per Student")
```











